---
title: "Best Practices colourblind-friendly"
format: html
---

## Colour-blind friendly colour palettes

First, create a matrix for the pie chart.

```{r}
x <- c(10,15,15,30,20,10)
```

If we put out the pie chart without any colour choices, the default colours will look like this:

```{r}
pie(x, main = "Pie Chart Default Colours")
```

Okabe & Ito (2008) proposed a colour palette that is accessible to people with different kinds of colour-blindness.[^1]

[^1]: https://easystats.github.io/see/reference/scale_color_okabeito.html

Luckily, there is a code, with which one can access this colour palette. It is named "Okabe-Ito".

```{r}
palette.colors(palette= "Okabe-Ito")
```

In the following code, I will manually assign the colours contained in "Okabe-Ito" to my colour palette that I wish to use later for the pie chart.

```{r}
okabe_ito_colours<- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00",  "#CC79A7","#999999")
```

Now I can use this palette for the pie chart. (rep(1,6) is used to choose 6 colours, since their are six pies in the chart, and okabe_ito_colours contains 9 colours.)

```{r}
pie(x, main = "Colour-blind friendly Pie Chart", rep(1, 6), col = okabe_ito_colours)
```

Now let's create a bar plot. First, create a x and y axis:

```{r}
x <- c("1", "2", "3", "4", "5")
y <- c(10, 3, 5, 2, 7)
barplot(y, names.arg = x, main = "Bar Plot no colours")
```

Without assigning colours, this type of plot results in a grey output.

Another colour-blind friendly palette is "R4". I will use the same procedure as before to create a colour-palette including R4 colours.

```{r}
palette.colors(palette= "R4")
R4_colours <- c("#000000","#DF536B", "#61D04F", "#2297E6", "#28E2E5", "#CD0BBC","#F5C710", "#9E9E9E")
```

Now, this colour palette will be included in our previously made bar plot.[^2]

[^2]: Sources:

    <https://www.w3schools.com/r/r_graph_bars.asp>

    <https://stackoverflow.com/questions/57153428/r-plot-color-combinations-that-are-colorblind-accessible>

```{r}
barplot(y, names.arg = x, main = "Colour-blind friendly Bar Plot", rep(1,5), col = R4_colours)
```
